Tidy data structure to support exploration and modeling of temporal data

### 2018-11-24 main.Rmd
added the linking of "key"

# Data semantics {#sec:semantics}

@@

The "key" specification contributes to creating a tsibble alongside the index. The concept of "key" is introduced to identify units or subjects that are observed over time in a data table. In the wide format, each column holds a series of values, so that the column itself serves for the sake of identification. In the long format, all column names are melt to the corresponding "key" values. But the "key" is much more flexible than simply column names. Because it is not constrained to a single field, but can be comprised of multiple fields. The identifying variables that the "key" is constituted of, remain the same as they are in the original table, with no additional manipulation.

Each tsibble must attach with a "key". It is normally a priori known by analysts. For example, Table \ref{tab:tb-sub} describes the number of tuberculosis cases for each gender across the countries every year. This data description suggests that columns `gender` and `country` have to be declared as the key. Lacking in either of two will be inadequate for the key and thus fail to construct a tsibble. The key is explicit when multiple units exist in the data. Key can be implicit when it finds a univariate series in the table, but it cannot be absent from a tsibble.

@@

Modeling is crucial to explanatory and predictive analytics, but often imposes stricter assumptions on tsibble data. The verbs listed in Table \ref{tab:??} ease the transition to a tsibble that suits modeling. A tidy forecasting framework built on top of tsibble is under development, which aims at promoting transparent forecasting practices and concise model representation. A tsibble usually contains multiple time series. Batch forecasting will be enabled if a univariate model, such as ARIMA and Exponential Smoothing, is applied to each time series independently. This yields a "mable" (short for model table), where each model only tags to each "key" value in tsibble to avoid expensive data copying and reduce model storage. The mable is further supplied to forecasting methods, to produce a tsibble in which each "key" along with its future time holds predictions. It also underlines the advantage of tsibble's "key" in linking between data inputs, models and forecasts. The modeling module will be fulfilled and integrated eventually.

@@

>>>>

# Contextual semantics {#sec:semantics}

@@

The "key" specification contributes to creating a tsibble alongside the index. The concept of "key" is introduced to identify units or subjects that are observed over time in a data table. In the wide format, each column holds a series of values, so that the column itself serves for the sake of identification. In the long format, all column names are melt to the corresponding "key" values. But the "key" is much more flexible than simply column names. Because it is not constrained to a single field, but can be comprised of multiple fields. The identifying variables that the "key" is constituted of, stay the same as they are in the original table, with no further tweaks.

Each tsibble must hold a "key". It is normally a priori known by analysts. For example, Table \ref{tab:tb-sub} describes the number of tuberculosis cases for each gender across the countries every year. This data description suggests that columns `gender` and `country` have to be declared as the key, similar to a panel variable for longitudinal data. Lacking in either of two will be inadequate for the key and thus fail to construct a tsibble. The key is explicit when multiple units exist in the data. Key can be implicit when it finds a univariate series in the table, but it cannot be absent from a tsibble.

Not only pinpoints a "key" observational units in the tsibble but also provides a solution to seamlessly link between the data, models, and forecasts. This neatly separates the data from models and forecasts, leaving more room for compulsory model components, such as coefficients, fitted values and residuals. More details are given in the following section.

@@

Modeling is crucial to explanatory and predictive analytics, but often imposes stricter assumptions on tsibble data. The verbs listed in Table \ref{tab:??} ease the transition to a tsibble that suits modeling. A tidy forecasting framework built on top of tsibble is under development, which aims at promoting transparent forecasting practices and concise model representation. A tsibble usually contains multiple time series. Batch forecasting will be enabled if a univariate model, such as ARIMA and Exponential Smoothing, is applied to each time series independently. This yields a "mable" (short for model table), where each model only tags to each "key" value in tsibble to avoid expensive data copying and reduce model storage. The mable is further supplied to forecasting methods, to produce a tsibble in which each "key" along with its future time holds predictions. It also underlines the advantage of tsibble's "key" in linking between data inputs, models and forecasts. Advanced forecasting techniques, such as vector autocorrelation, hierarchical reconciliation, and ensembles, can be developed in alike spirit. The modeling module will be fulfilled and integrated eventually.

@@

# Software structure

* Extensibility
